Understanding LIME

### 2019-09-06 README.Rmd
started adding some code that demonstrates how the internal lime functions work

- Graphics Group
    - email list
    - what should we do about gganimate/Halloween?

@@

- code up Laugel et al's work
- read through papers giving an overview of the current explanation methods
- check to see if there a paper showing how ridge regression is affected by sample size changes
- try fitting the ridge regression with the observed data only to see how it compares to the model fit with the ridge regression

>>>>

- Graphics Group
    - email list
    - what should we do about gganimate/Halloween?
- LIME
    - I made a mistake when I told you that the full test dataset gets included in the ridge regression model (only the case of interest in included)

@@

- finish going through code with the explain function
- write procedure section of LIME paper
- draft the rest of the paper
- read through papers giving an overview of the current explanation methods

### 2019-09-06 journals/02-understanding_lime/02-understanding_lime.Rmd
started adding some code that demonstrates how the internal lime functions work

@@

8. Obtains predictions for all of the permutated data points based on the original model using the [`predict_model`](#predict_model) function.

@@

@@

2. Create the perturbations based on the type of variable and using the feature distribution determined by the lime function. The cases are as follows.

@@

    - **integer variable**: round the permutations to the nearest integer

            if (is.integer(cases[[i]])) {
              as.integer(round(perms))
            }

    - **other**:

            } else {
              perms
            }
            }), stringsAsFactors = FALSE)

3. Give the permutations the same name as the variables

@@

4. Add the cases back into the data frame of the permutations

@@

5. Return the permutations.

@@

@@

@@

>>>>

In order to run the *lime* functions, here I load the bullet data to apply the code to.

```{r}
# Load in the bullet training and testing data
hamby173and252_train <- read.csv("../../../data/hamby173and252_train.csv")
hamby224_test <- read.csv("../../../data/hamby224_test.csv")

# Obtain features used when fitting the rtrees random forest
rf_features <- rownames(bulletr::rtrees$importance)
```

@@

8. Obtains predictions for all of the permutated data points based on the original model using the [`predict_model`](#predict_model) function. Use the function [`set_labels`](#set_labels.)

@@

#### Implementation

```{r}
library(randomForest)
x <- hamby224_test %>% 
  arrange(case) %>%
  select(rf_features) %>% 
  na.omit()
explainer <- lime:::lime.data.frame(x = hamby173and252_train %>%
                                      select(rf_features), 
                                    y = hamby173and252_train %>%
                                      select(samesource), 
                                    model = as_classifier(bulletr::rtrees))
label = TRUE
n_labels = NULL
n_features = 3
n_permutations = 5000
feature_select = 'auto'
dist_fun = "gower"
kernel_width = NULL
gower_pow = 1
```

```{r}
assertthat::assert_that(lime:::is.data_frame_explainer(explainer))
```

```{r}
#m_type <- lime:::model_type(explainer)
o_type <- lime:::output_type(explainer)
```

```{r}
assert_that(is.null(labels) + is.null(n_labels) == 1, 
   msg = "You need to choose between labels and n_labels parameters.")
```

```{r}
assert_that(is.count(n_features))
assert_that(is.count(n_permutations))
```

```{r}
if (is.null(kernel_width)) {
   kernel_width <- sqrt(ncol(x)) * 0.75
}
kernel <- lime:::exp_kernel(kernel_width)
```

```{r}
case_perm <- lime:::permute_cases.data.frame(x,
                                             n_permutations,
                                             explainer$feature_distribution,
                                             explainer$bin_continuous,
                                             explainer$bin_cuts,
                                             explainer$use_density)
head(case_perm)
```

```{r}
case_res <- lime:::predict_model(explainer$model, 
                                 explainer$preprocess(case_perm),
                                 type = o_type)
case_res <- lime:::set_labels(case_res, explainer$model)
head(case_res)
```

```{r}
case_ind <- split(seq_len(nrow(case_perm)), rep(seq_len(nrow(x)), each = n_permutations))
case_ind[[1]][1:10]
case_ind[[2]][1:10]
case_ind[[3]][1:10]
```

```{r}
str(seq_along(case_ind))
```

```{r}
i <- case_ind[[1]]
str(i)
```

```{r}
if (dist_fun == "gower") {
  sim <- 1 - (gower_dist(case_perm[i[1], , drop = FALSE], 
    case_perm[i, , drop = FALSE]))^gower_pow
}
str(sim)
```

```{r}
perms <- lime:::numerify(case_perm[i, ], 
                         explainer$feature_type, 
                         explainer$bin_continuous,
                         explainer$bin_cuts)
head(perms)
```

```{r}
# error?
# res <- lime:::model_permutations(as.matrix(perms), 
#                                  case_res[i, , drop = FALSE], 
#                                  sim, 
#                                  labels, 
#                                  n_labels, 
#                                  n_features, 
#                                  feature_select)
```

@@

2. Create the perturbations based on the type of variable and using the feature distribution determined by the lime function. The cases are as follows. The function does this process one variable at a time where a case represents on variable.

@@

3. If the variable in the original dataset is an integer, then convert the perturbations to integers.

        if (is.integer(cases[[i]])) {
          as.integer(round(perms))
        } else {
          perms
        }

4. Give the permutations the same name as the variables.

@@

5. Add the cases back into the data frame of the permutations, so that the case of interest gets inserted at the beginning of each set of perturbations.

@@

6. Return the permutations.

@@

The final `perm` will look like...

| Case | Obs | 
| ---- | --------- | 
| 1 | 1st case of interest | 
| 1 | perturbation 1,1 | 
| 1 | perturbation 1,2 |
| ... | ... | 
| 1 | perturbation 1,m |
| 2 | 2nd case of interest | 
| 2 | perturbation 2,1 | 
| 2 | perturbation 2,2 |
| ... | ... | 
| 2 | perturbation 2,m |
| ... | ... | 
| n | nth case of interest | 
| n | perturbation n,1 | 
| n | perturbation n,2 |
| ... | ... | 
| n | perturbation n,m |

@@

#### Implementation

```{r}
cases = x
n_permutations = n_permutations
feature_distribution = explainer$feature_distribution
bin_continuous = explainer$bin_continuous
bin_cuts = explainer$bin_cuts
use_density = explainer$use_density
```

```{r}
nrows <- nrow(cases) * n_permutations
```

```{r}
seq <- seq_along(cases)
seq
```

```{r}
i <- seq[1]
# Case when is.numeric(cases[[i]]) && bin_continuous
bin <- sample(seq_along(feature_distribution[[i]]), nrows, TRUE,
              as.numeric(feature_distribution[[i]]))
head(bin)
perms <- diff(bin_cuts[[i]])[bin] * runif(nrows) + bin_cuts[[i]][bin]
head(perms)
```

```{r}
if (is.integer(cases[[i]])) {
    perms <- as.integer(round(perms))
  } else {
    perms <- perms
  }
head(perms)
```

```{r}
perm <- as.data.frame(lapply(seq_along(cases), function(i) {
         perms <- if (is.numeric(cases[[i]]) && bin_continuous) {
             bin <- sample(seq_along(feature_distribution[[i]]), 
                 nrows, TRUE, as.numeric(feature_distribution[[i]]))
             diff(bin_cuts[[i]])[bin] * runif(nrows) + bin_cuts[[i]][bin]
         }
         if (is.integer(cases[[i]])) {
             as.integer(round(perms))
         }
         else {
             perms
         }
}), stringsAsFactors = FALSE)
names(perm) <- names(cases)
head(perm)
```

```{r}
# The first row gets replaced with the case of interest for each set of perturbations
perm[seq.int(1, by = n_permutations, length.out = nrow(cases)), ] <- cases
head(perm)
```

@@

### `set_labels` {#set_labels .tabset}

- `res` = the predictions (`case_res`)
- `model` = `explainer$model`

```{r}
# Print the function (and check to see if there have been any changes)
func_check("set_labels", "0.5.0")
```
